---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NearestBalance

<!-- badges: start -->
<!-- badges: end -->

The goal of NearestBalance is to ...

## Installation

You can install the released version of NearestBalance from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("NearestBalance")
```

## Quick start guide

Load data for the example

``` {r example, warning=FALSE, verbose=FALSE}
library(NearestBalance)
library(zCompositions)
library(selbal)
test_data <- selbal::HIV[1:60]
abundance <- cmultRepl(test_data)
```


## Principal balance analysis with NearestBalance
``` {r PBA}
nb_1 <- nb_pca(abundance)
plot_nb_pca(nb_1, colour = HIV$MSM, pch = HIV$HIV_Status)

# first principal balance
nb_1$nb$b1
```
## Regression analysis
``` {r lm}
nb_2 <- nb_lm(abundance = abundance,
              metadata = HIV,
              pred = "MSM",
              cov = "HIV_Status")
```

Statistical significance of the association
```{r signif}
summary(manova(nb_2$lm_res))
```

nearest balance interpretation of the regression coefficient
```{r lm balance}
nb_2$nb$b1
```

## Interpretation of the SVM results
``` {r svm}
nb_3 <- nb_svm(abundance = abundance,
               f = HIV$HIV_Status)
```

The nearest balance for the discriminating direction
``` {r}
nb_3$nb$b1
```

## Interpretation of the LDA results
``` {r lda}
nb_4 <- nb_lda(abundance = abundance,
               f = HIV$HIV_Status)
```

The nearest balance for the discriminating direction
``` {r}
nb_4$nb$b1
```

## Interpretation of differences between two samples
```{r}
nb_5 <- nb_shift(abundance = abundance[,1:15],
                 samp_1 = rownames(abundance)[1],
                 samp_2 = rownames(abundance)[2],
                 type = "two_balances")
```

The nearest balance
```{r}
nb_5$b1
```

Impacts of the balances
```{r}
nb_5$impactsdthf
```

