---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


## Installation

Install dependences if needed:
``` {r}
cran_pckgs <- c("data.tree", "data.table", "stringr",
                "e1071", "ggplot2", "partitions", "devtools",
                "reshape2", "forcats", "ggh4x")
cran_pckgs <- setdiff(cran_pckgs, installed.packages())

if (length(cran_pckgs) >0){
  install.packages(cran_pckgs)
}

if (! "balance" %in% installed.packages()){
  devtools::install_github("tpq/balance")
}
```
Install the package:
``` r 
devtools::install_bitbucket("knomics/nearestbalancedev")  
```
## Quick start guide
Please, install several packages to run the examples:
``` {r installstion_for_example, warning=FALSE, verbose=FALSE}
pckgs_for_example <- setdiff(c("selbal", "zCompositions"),installed.packages())

if (length(pckgs_for_example) > 0){
  install.packages(pckgs_for_example)  
}
```

Load data for the example

``` {r example, warning=FALSE, verbose=FALSE}
library(NearestBalance)
library(zCompositions)
library(selbal)
test_data <- selbal::HIV[1:60]
abundance <- cmultRepl(test_data)
```


## Principal balance analysis with NearestBalance
``` {r PBA, echo = FALSE}
nb_1 <- nb_pca(abundance)
plot_nb_pca(nb_1, colour = HIV$MSM, pch = HIV$HIV_Status)

# first principal balance
nb_1$nb$b1
```
## Regression analysis
``` {r lm}
nb_2 <- nb_lm(abundance = abundance,
              metadata = HIV,
              pred = "MSM",
              cov = "HIV_Status")
```

Statistical significance of the association
```{r signif}
summary(manova(nb_2$lm_res))
```

nearest balance interpretation of the regression coefficient
```{r lm balance}
nb_2$nb$b1
```

Heatmap:
```{r lm heatmap, fig.height = 9, fig.width = 12}
heatmap_with_split(abundance = abundance,
                   metadata = HIV, 
                   formula = ~ MSM + HIV_Status, 
                   balance = nb_2$nb$b1,
                   show_samp_names = F,
                   num_name = "taxa_MSM",
                   den_name = "taxa_nonMSM",
                   others_name = "not assosiated with MSM")
```

## Interpretation of the SVM results
``` {r svm}
nb_3 <- nb_svm(abundance = abundance,
               f = HIV$HIV_Status)
```

The nearest balance for the discriminating direction
``` {r}
nb_3$nb$b1
```

## Interpretation of the LDA results
``` {r lda}
nb_4 <- nb_lda(abundance = abundance,
               f = HIV$HIV_Status)
```

The nearest balance for the discriminating direction
``` {r}
nb_4$nb$b1
```

## Interpretation of differences between two samples
```{r}
counts <- HFD[1:193]
counts_filt <- counts[, colSums(counts > 0) > 0.5*nrow(counts)]
abundance <- cmultRepl(counts_filt)

meta <- HFD[, c("sample_id", "subject_id", "group")]
pairs <- reshape2::dcast(meta, subject_id ~ group, value.var = "sample_id")

nb_5 <- nb_shift(abundance = abundance,
                 samp_1 = pairs$after[1],
                 samp_2 = pairs$before[1])
```

The nearest balance to the microbiome shift of the first subject
```{r}
nb_5$b1
```


Impact of the nearest balance in the total shift
```{r}
nb_5$impact
```


## Interpretation of the mean shift in the sample
```{r}
nb <- nb_mean_shift(abundance, pairs$before, pairs$after, "two_balances")

plot(nb$coord$b1, nb$coord$b2, col = as.factor(HFD$group), xlab = "b1", ylab = "b2")
arrows(mean(nb$coord[pairs$before, "b1"]),
  mean(nb$coord[pairs$before, "b2"]),
  mean(nb$coord[pairs$before, "b1"]) + nb$nb$coord["b1"],
  mean(nb$coord[pairs$before, "b2"]) + nb$nb$coord["b2"],
  lwd =3)
```

Statistical significance of the shift
```{r}
summary(manova(nb$lm_res),intercept = T)
```

Members of the nearest balance to the mean shift
```{r}
nb$nb$b1
```
Impact
```{r}
nb$nb$impacts
```
Proportion of noise in data to the mean shift
```{r}
nb$noise
```





